### 2023-01-20 output/tables/ddiv_show.tex
DDIV

\begin{table}[htbp]\centering
\begin{flushright}
\caption{Replication of Panel A in Table 3 of Duflo (2001)}
\resizebox{0.9\linewidth}{!}{
\begin{tabular}{l*{6}{c}} \hline \\ \hline
& \multicolumn{3}{c}{College enrollment} & \multicolumn{3}{c}{Fam inc}\\ \cline{2-4} \cline{5-7}
& \multicolumn{3}{c}{Level of drug arrests} & \multicolumn{3}{c}{Level of drug arrests}\\ \cline{2-4} \cline{5-7}
& High & Low & Diff. & High & Low & Diff. \\
& (1) & (2) & (3) & (4) & (5) & (6) \\ \hline
\input{ddiv.tex} \\ \hline
\multicolumn{4}{l}{\footnotesize Clustered (state-year) robust standard errors in parentheses}\\
\multicolumn{4}{l}{\footnotesize CPS education supplement weights used. Males only}\\
\end{tabular}
}
\end{flushright}
\end{table}