### 2023-03-25 source/paper/tables.tex

\import{../../output/tables/summ_stats/}{cps_summ_stats.tex}

\begin{footnotesize}
  \noindent Note: Sample means (with education weights) are displayed from the CPS October supplement merged with the UCR dataset from 1984-1992 and 2000-2016. The sample in columns 1 and 2 are defined as persons aged 18-24 in 1986, and the sample in columns 3 and 4 are defined as persons aged 18-24 in 2010, who were not incarcerated at the time of the survey. This table is partially replicated from \cite{britton2022}.
\end{footnotesize}

\import{../../output/tables/summ_stats/}{ucr_summ_stats.tex}

% Britton Table 2
\import{../../output/tables/}{britton_table2_DiD.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black and white males aged 18-24 in 1986 who were not incarcerated at the time of the survey.
  This table is partially replicated from \cite{britton2022}.
\end{footnotesize}

\import{../../output/tables/}{britton_table2_DiD_control_experiment.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black and white males aged 30-50 in 1986 who were not incarcerated at the time of the survey.
  This table is a control experiment for table 2 in \cite{britton2022}.
\end{footnotesize}
\clearpage

\import{../../output/tables/}{britton_table3_DiD.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black males and black females aged 18-24 in 1986 who were not incarcerated at the time of the survey.
  This table is partially replicated from \cite{britton2022}.
\end{footnotesize}

\import{../../output/tables/}{britton_table3_DiD_control_experiment.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black males and black females aged 30-50 in 1986 who were not incarcerated at the time of the survey.
  This table is a control experiment for table 3 in \cite{britton2022}.
\end{footnotesize}
\clearpage

\import{../../output/tables/}{fair_sentencing_DiD_t1.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black and white males aged 18-24 in 2010 who were not incarcerated at the time of the survey.
\end{footnotesize}

\import{../../output/tables/}{fair_sentencing_DiD_t1_control_experiment.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black and white males aged 30-50 in 2010 who were not incarcerated at the time of the survey.
\end{footnotesize}
\clearpage

\import{../../output/tables/}{fair_sentencing_DiD_t2.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black males and females aged 18-24 in 2010 who were not incarcerated at the time of the survey.
\end{footnotesize}

\import{../../output/tables/}{fair_sentencing_DiD_t2_control_experiment.tex}
\begin{footnotesize}
  \noindent Note: Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as black males and females aged 30-50 in 2010 who were not incarcerated at the time of the survey.
\end{footnotesize}
\clearpage

\import{../../output/tables/}{DiD_1986_high_low.tex}
\begin{footnotesize}
  \noindent Note: Treated observations are defined as those living in states with a high-drug arrest rate for black adults, where high black adult drug arrest states are defined to be those above the 75th percentile in 1984. Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as males aged 18-24 in 1986 who were not incarcerated at the time of the survey.
\end{footnotesize}

\import{../../output/tables/}{DiD_1986_high_low_cont.tex}
\begin{footnotesize}
  \noindent Note: Treatment is continuous. Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as males aged 18-24 in 1986 who were not incarcerated at the time of the survey.
\end{footnotesize}
\clearpage

\import{../../output/tables/}{DiD_1986_high_low_jb.tex}
\begin{footnotesize}
  \noindent Note: Treated observations are defined as those living in states with a high-drug arrest rate for black juveniles, where high black juvenile drug arrest states are defined to be those above the 75th percentile in 1984. Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as males aged 18-24 in 1986 who were not incarcerated at the time of the survey.
\end{footnotesize}
\import{../../output/tables/}{DiD_1986_high_low_jb_cont.tex}
\clearpage

\import{../../output/tables/}{DiD_1986_high_low_jb_cont_control_experiment_female.tex}
\import{../../output/tables/}{DiD_1986_high_low_jb_cont_control_experiment_old.tex}

\clearpage

\import{../../output/tables/}{DiD_2010_high_low.tex}
\begin{footnotesize}
  \noindent Note: Treated observations are defined as those living in states with a high-drug arrest rate for black adults, where high black adult drug arrest states are defined to be those above the 75th percentile in 2008. Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as males aged 18-24 in 1986 who were not incarcerated at the time of the survey.
\end{footnotesize}

\import{../../output/tables/}{DiD_2010_high_low_cont.tex}
\begin{footnotesize}
  \noindent Note: Treatment is continuous. Estimates are weighted using CPS October supplement weights. Robust standard errors are clustered at the state level. Controls: age, age-squared, Latino ethnicity, yearly state average unemployment rates, and (binned) family income. The sample is defined as males aged 18-24 in 2010 who were not incarcerated at the time of the survey.
\end{footnotesize}
\clearpage

\import{../../output/tables/}{ddd_1986.tex}

\import{../../output/tables/}{ddiv_show.tex}