### 2023-04-18 output/tables/ddiv/ddiv_show2.tex
Update

\begin{table}[htbp]
\centering
\caption{Mean of Drug-related Arrest Rate and College Enrollment Rate By Time and High/low Drug Arrest State}
\resizebox{0.9\linewidth}{!}{
\begin{tabular}{l*{6}{c}} 
\hline
\hline
& \multicolumn{3}{c}{Drug-Related Arrest Rate} & \multicolumn{3}{c}{College Enrollment} \\ 
\cline{2-4} \cline{5-7}
& \multicolumn{3}{c}{Level of drug arrests} & \multicolumn{3}{c}{Level of drug arrests} \\ 
\cline{2-4} \cline{5-7}
& High & Low & Diff. & High & Low & Diff. \\
& (1) & (2) & (3) & (4) & (5) & (6) \\ 
\hline
\input{ddiv_high_low.tex} \\ 
\hline

\end{tabular}
}
\end{table}